---
title: "Homework 1"
author: "Aaron Banlao"
output: pdf_document
---


```{r}
library(fpp3)
library(tsibbledata)
library(tsibble)
library(ggplot2)
library(dplyr)
```

# Problem 1

Explore the following four time series: Bricks from aus_production, Lynx from pelt, Close from gafa_stock, Demand from vic_elec.

Use ? (or help()) to find out about the data in each series.
What is the time interval of each series?
Use autoplot() to produce a time plot of each series.
For the last plot, modify the axis labels and title.

## a)

### Code) 
```{r}
head(gafa_stock)
autoplot(gafa_stock, Close) + labs(title = "Google Amazon Facebook Apple Stock")
```


```{r}
head(PBS)

PBS %>% 
  summarise(Total = sum(Cost)) %>% 
  autoplot(Total) + labs(title = "Monthly Aggregated Medical Australia Prescription Cost")
```
```{r}
head(vic_elec)
autoplot(vic_elec, Temperature) + labs(title =  "Electricity Demand for Victoria in regards of Temperature", x = "Per Half Hour")
```

```{r}
head(pelt)
autoplot(pelt, Hare) + labs(title = "Number of Snowshoe Hare Pelts Traded")
```

## b)

### Answer)

1) The time interval for gafa_stock is daily (business trading days)
2) The time interval for PBS is monthly
3) The time interval for vic_elec is every 30 minutes
4) The time interval for pelt is yearly 

# Problem 2 

The USgas package contains data on the demand for natural gas in the US.

Install the USgas package.
Create a tsibble from us_total with year as the index and state as the key.
Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).

## a)
```{r}
#install.packages("USgas")
library(USgas)
```

## b)
```{r}
head(us_total)
```
```{r}
total <- us_total %>% 
  filter(state == c("Maine", "Vermont", "New Hampshire", "Massachusetts", "Connecticut", "Rhode Island")) %>% 
  mutate(y = y/1000) %>% 
  as_tsibble(key = state, index = year)
```

## c)
```{r}
autoplot(total) + labs(y = "Gas Consumption", title = "Natural Gas Consumption by State")
```


# Problem 3 

Monthly Australian retail data is provided in aus_retail. Select one of the time series as follows (but choose your own seed value):

set.seed(12345678)
myseries <- aus_retail |>
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))
Explore your chosen retail time series using the following functions:

autoplot(), gg_season(), gg_subseries(), gg_lag(),

ACF() |> autoplot()

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

## Answer)

Japan departures have an increasing trend up until the late 1990's and then has a decreasing trend from there on, compared to the other countries, where their departures are increasing over time. In terms of seasonality, Japan has an alternating increasing and decreasing trend between quarters.


## Code)

```{r}
head(aus_arrivals)
autoplot(aus_arrivals)
```

```{r}
gg_season(aus_arrivals)
```

```{r}
gg_subseries(aus_arrivals)
```



# Problem 4

Use the following graphics functions: autoplot(), gg_season(), gg_subseries(), gg_lag(), ACF() and explore features from the following time series: “Total Private” Employed from us_employment, Bricks from aus_production, Hare from pelt, “H02” Cost from PBS, and Barrels from us_gasoline.

Can you spot any seasonality, cyclicity and trend?
What do you learn about the series?
What can you say about the seasonal patterns?
Can you identify any unusual years?

```{r}
head(aus_retail)
```

```{r}
set.seed(101)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))
```

```{r}
autoplot(myseries)
```

There is an apparent upward trend over time. There also seems to be a cyclical pattern but the fluctuations are greater after the 2000's.

```{r}
gg_season(myseries)
```

There is a seasonal pattern going on with sales jumping in December and then dropping for January. Sales are steady until September where they begin to increase until December

```{r}
gg_subseries(myseries)
```

```{r}
gg_lag(myseries)
```