---
title: "Homework 3"
author: "Aaron Banlao"
output:
  html_document:
    df_print: paged
---



```{r}
library(tidyverse)
library(fpp3)
library(tsibble)
library(tsibbledata)
library(feasts)
library(broom)
library(fracdiff)
library(fable)
library(glue)
```


# Problem 1

Write a function to compute the mean and standard deviation of a time series, and apply it to the PBS data. Plot the series with the highest mean, and the series with the lowest standard deviation.

## Code

```{r}
head(PBS)
```


```{r}
maxmean_PBS <- PBS %>% 
  features(Cost, list(mean = mean)) %>% 
  filter(mean == max(mean))

maxmean_PBS
```

```{r}
PBS %>% 
  filter(Concession == "Concessional",
         Type == "Co-payments", 
         ATC1 == "C",
         ATC2 == "C10") %>% 
  autoplot() +
  labs(title = "Highest Mean Series")
```


```{r}
PBS %>% 
  features(Cost, list(sd = sd)) %>% 
  filter(sd == min(sd))
```
```{r}
PBS %>% 
  filter(Concession == "General", 
         Type == "Co-payments",
         ATC1 == "S", 
         ATC2 == "S") %>% 
  autoplot() +
  labs(title = "Lowest Standard Deviation Series")
```

# Problem 3)

Use a feature-based approach to look for outlying series in the PBS data. What is unusual about the series you identify as “outliers”.

## Answer)

The outliers we have picked out have an unusual "seasonal" pattern, or lack of seasonality pattern we have observed in other datasets. These four outliers also show no trend. 

## Code)

```{r}
pbs_features <- PBS %>% 
  features(Cost, feature_set(pkgs = "feasts")) %>% 
  select(-...26) %>% 
  na.omit()
```

```{r}
head(pbs_features)
```



```{r}
pbs1 <- pbs_features %>% 
  select(-Concession, -Type, -ATC1, -ATC2) %>% 
  prcomp(scale = T) %>% 
  augment(pbs_features)
```

```{r}
pbs1 %>% 
  ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Type)) +
  geom_point()
```

```{r}
outliers <- pbs1 %>% 
  filter(.fittedPC1 > 6) %>% 
  select(Concession, Type, ATC1, ATC2)

outliers
```

```{r}
outliers %>% 
  left_join(PBS, by = c("Concession", "Type", "ATC1", "ATC2")) %>% 
  mutate(Series = glue("{Concession}", "{Type}", "{ATC1}", "{ATC2}", .sep = "\n\n")) %>% 
  ggplot(aes(x = Month, y = Cost)) +
  geom_line() +
  facet_grid(Series ~ ., scales = "free") + 
  labs(title = "Outliers")
```