---
title: "Project"
author: "Aaron Banlao"
output: pdf_document
---


Install the COVID19 Data Hub R package. Select at least two countries to work with of interest to you and make forecasts of the number of cases for the next 12 months.

Build the following models:

Build a linear regression model using TSLM().
Build an appropriate exponential smoothing model.


```{r}
library(pacman)
p_load(fpp3, tidyverse, doParallel)
```


```{r}
covid <- read.csv("1.csv", header = T)
```

```{r}
head(covid)
```

```{r}
covid$id[covid$id == "59a13ceb"] <- "United States"
covid$id[covid$id == "78833522"] <- "China"
```

```{r}
covid <- covid %>% 
  filter(id == "United States" | id == "China")
head(covid)
```

```{r}
covid$date <- as.Date(covid$date)
```




```{r}
covid <- covid %>% 
  as_tsibble(key = "id", index = "date")
head(covid)
```

```{r}
class(covid)
```


```{r}
covid <- covid %>% 
  dplyr::select(id, date, confirmed)
```

```{r}
covid <- covid %>% 
  drop_na()
```


# Time Series Linear Model

## China

```{r}
china <- covid %>% 
  filter(id == "China") %>% 
  select(date, confirmed)

head(china)
```



```{r}
china_fit <- china %>% 
  model(TSLM(confirmed ~ trend() + season()))

china_fit %>% 
  forecast(h = 365) %>% 
  autoplot(china) +
  labs(title = "China Confirmed Cases Forecast")
```

## United States

```{r}
us <- covid %>% 
  filter(id == "United States") %>% 
  select(date, confirmed)

head(us)
```


```{r}
us_fit <- us %>% 
  model(TSLM(confirmed ~ trend() + season()))

us_fit %>% 
  forecast(h = 365) %>% 
  autoplot(us) +
  labs(title = "US Confirmed Cases Forecast")
```


# Exponential Smoothing 

## China

### Best Model

According to the RSME value, the appropriate exponential model is the holts method.

```{r}
registerDoParallel(cores = 16)
china %>% 
  stretch_tsibble() %>% 
  model(dholts = ETS(confirmed ~ error("A") + trend("Ad") + season("N")),
        holts = ETS(confirmed ~ error("A") + trend("A") + season("N")),
        ses = ETS(confirmed ~ error("A") + trend("N") + season("N"))) %>% 
  forecast(h = 365) %>% 
  accuracy(china)
```

```{r}
china %>% 
  model(holts = ETS(confirmed ~ error("A") + trend("A") + season("N"))) %>% 
  forecast(h = 365) %>% 
  autoplot(china) +
  labs(title = "Predicted China Confirmed Cases (Holts)")
```


## US

### Best Model

According to the RSME values, the appropriate exponential model is the damped holts method.

```{r}
us %>% 
  stretch_tsibble(.init = 10) %>% 
  model(dholts = ETS(confirmed ~ error("A") + trend("Ad") + season("N")),
        holts = ETS(confirmed ~ error("A") + trend("A") + season("N")),
        ses = ETS(confirmed ~ error("A") + trend("N") + season("N"))) %>% 
  forecast(h = 365) %>% 
  accuracy(us)
```

```{r}
us %>% 
  model(dholts = ETS(confirmed ~ error("A") + trend("Ad") + season("N"))) %>% 
  forecast(h = 365) %>% 
  autoplot(us) +
  labs(title = "Predicted US Confirmed Cases (Damped Holts)")

```